### 2025-01-01 acronyms.tex
Updates for 2024

\acrodef{AEA}{American Economic Association}
\acrodef{AI}{artificial intelligence}
\acrodef{AJPS}{American Journal of Political Science}
\acrodef{BPLIM}{Banco de Portugal Microdata Research Laboratory}
\acrodef{BLS}{Bureau of Labor Statistics}
\acrodef{CIQSS}{Centre interuniversitaire québécois de statistiques sociales}
\acrodef{CEA}{Canadian Economics Association}
\acrodef{CJE}{Canadian Journal of Economics}
\acrodef{CSWEP}{AEA Committee on the Status of Women in the Economics Profession}
\acrodef{DCAP}{data and code availability policy}
\acrodef{DCAS}{Data and Code Availability Standard}
\acrodef{DOI}{Digital Object Identifier}
\acrodef{EEA}{European Economics Association}
\acrodef{EJ}{Economic Journal}
\acrodef{EI}{Economic Inquiry}
\acrodef{ERS}{Economic Research Service}
\acrodef{FAIR}{Findable, Accessible, Interoperable, Re-usable}
\acrodef{FAQ}{frequently asked questions}
\acrodef{FSRDC}{Federal Statistical Research Data Centers}
\acrodef{GPU}{graphical processing unit}
\acrodef{HRS}{Health and Retirement Study}
\acrodef{IAB}{Research Data Center (FDZ) at the Institute for Employment Research}
\acrodef{IRS}{Internal Revenue Service}
\acrodef{ICPSR}{Inter-university Consortium for Political and Social Research}
\acrodef{JASA}{Journal of the American Statistical Association}
\acrodef{JEEA}{Journal of the European Economic Association}
\acrodef{JFE}{Journal of Financial Economics}
\acrodef{JPC}{Journal of Privacy and Confidentiality}
\acrodef{HDSR}{Harvard Data Science Review}
\acrodef{SPCR}{Society for Confidentiality and Privacy Research}
\acrodef{LLM}{large language models}
\acrodef{LMIC}{low- and middle-income countries}
\acrodef{NACJD}{National Archive of Criminal Justice Data}
\acrodef{NBER}{National Bureau of Economic Research}
\acrodef{OLDA}{Ohio Longitudinal Data Archive}
\acrodef{PAP}{pre-analysis plans}
\acrodef{PII}{personally identifiable information}
\acrodef{PSID}{Panel Study of Income Dynamics}
\acrodef{RAM}{random access memory}
\acrodef{RCT}{randomized control trial}
\acrodef{ReStud}{Review of Economic Studies}
\acrodef{TIER}{Project TIER (Teaching Integrity in Empirical Research)}
\acrodef{USDA}{{US} Department of Agriculture}
\acrodef{WEAI}{Western Economics Association International}
\newcommand{\aeadcr}{AEA Data and Code Repository}
\newcommand{\dcap}{Data and Availability Policy}
\newcommand{\rctr}{AEA RCT Registry}

\acrodef{AER}{American Economic Review}
\acrodef{AERI}{American Economic Review: Insights}
\acrodef{AEJAPP}{American Economic Journal: Applied Economics}
\acrodef{AEJPOL}{American Economic Journal: Economic Policy}