### 2025-02-07 outputs/table_stata-licenses.tex

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Fri Feb  7 16:33:05 2025
\begin{table}[h]
\centering
\caption{Software licensing internationally} 
\begin{tabular}{lrrr}
  \toprule
Country & Average monthly income & Stata license & Percent \\ 
  \midrule
USA & 6704 & 690 & 10 \\ 
  Greece & 1883 & 690 & 37 \\ 
  Vietnam & 343 & 220 & 64 \\ 
   \bottomrule
\end{tabular}
\end{table}