### 2025-02-08 outputs/table_stata-licenses.tex
Added better formatting

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Fri Feb  7 20:38:40 2025
\begin{table}[h]
\centering
\caption{Software licensing internationally} 
\begin{tabular}{llll}
  \toprule
Country & Average.monthly.income & Stata.license & Percent \\ 
  \midrule
USA & 6,704 & 690 & 10.3 \\ 
  Greece & 1,883 & 690 & 36.6 \\ 
  Vietnam & 343 & 220 & 64.1 \\ 
   \midrule \multicolumn{4}{l}{\textit{Notes:} Amounts in USD as of 2025. Source for Stata license prices: Stata website.} 
 \bottomrule
\end{tabular}
\end{table}