### 2025-02-09 outputs/table_restrictions.tex
Updated labeling to be tex-safe

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Sat Feb  8 19:16:13 2025
\begin{table}[h]
\centering
\caption{Example restrictions} 
\label{tab:restrictions}
\begin{tabular}{llp{4in}}
  \toprule
Dataset & Variant & Usage.restrictions.and.justification \\ 
  \midrule
SOEP & European version & No redistribution and access by unauthorized individuals. Compliance with  German Federal Data Protection Act. \\ 
  SOEP & International version & Same as European version, but 95\% sample, and institutional agreement required. \\ 
  DHS &  & No redistribution and access by unauthorized individuals. To ensure that provider meets host-country agreements, such as compliance with usage by  real people with legitimate research purposes. \\ 
  WVS & Except for joint WVS-EVS files & The data can be used freely for non-commercial purposes such as research, publication, teaching. Data redistribution is prohibited: publication of the original WVS datasets at other online platforms is against the WVSA Constitution. \\ 
  PSID & Public data & Use the data in the PSID data sets only for scientific research and aggregate statistical reporting; Make no attempts to identify study participants; \\ 
  PSID & Restricted data & In order to safeguard the confidentiality of respondents at the highest level, some data are provided only under conditions of a restricted use contract, including human subjects review and data security plan. \\ 
   \bottomrule
\end{tabular}
\end{table}