### 2025-03-05 outputs/table_access_provision.tex
Adding table of access

\begin{table}[!h]
\centering
\caption{Access categories and whether data can be shared privately}
\centering
\begin{tabular}[t]{lrllr}
\toprule
Access restrictions & No & Yes & Percent & Total\\
\midrule
No & 50 & n/a & (0.00\%) & 50\\
Very Easy to Obtain & 11 & 7 & (38.89\%) & 18\\
Moderately Easy to Obtain & 3 & 5 & (62.50\%) & 8\\
Moderately Difficult to Obtain & 16 & 9 & (36.00\%) & 25\\
Very Difficult to Obtain & 9 & 5 & (35.71\%) & 14\\
\addlinespace
Total & 89 & 26 & (22.61\%) & 115\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{*} Percentages are calculated as the number of 'Yes' divided by the}\\
\multicolumn{5}{l}{total number of responses. An article can have multiple categories}\\
\multicolumn{5}{l}{of data; the sum of responses is therefore higher than the number}\\
\multicolumn{5}{l}{of articles.}\\
\end{tabular}
\end{table}