### 2025-03-06 outputs/table_categories.tex
Adding categories

% latex table generated in R 4.4.2 by xtable 1.8-4 package
% Thu Mar  6 03:27:42 2025
\begin{table}[h]
\centering
\caption{Restriction categories} 
\label{tab:categories}
\begin{tabular}{lp{5in}}
  \toprule
Category & Explanation \\ 
  \midrule
Very Easy to Obtain & Request takes just a few minutes with no associated costs and the expected response is within a few days \\ 
  Moderately Easy to Obtain & Request takes less than an hour with minimal cost \\ 
  Moderately Difficult to Obtain & Request requires a multipage application; request needs university approvals; request involves significant cost; there is some uncertainty as to whether the proposal will be granted \\ 
  Very Difficult to Obtain & Request must be made in person and/or access is provided only in person; request requires substantial funding; data and/or access mechanism may no longer exist \\ 
   \bottomrule
\end{tabular}
\end{table}