### 2025-03-06 outputs/table_access_provision.tex
Adding spacing

\begin{table}[!h]
\centering
\caption{\label{tab:access_provision}Access categories and whether data can be shared privately}
\centering
\begin{tabular}[t]{lllrl}
\toprule
Access restrictions & No & Yes & Total & Percent\\
\midrule
No &  & n/a & 50 & \\
\midrule\\
Very Easy to Obtain & 11 & 7 & 18 & (38.89\%)\\
Moderately Easy to Obtain & 3 & 5 & 8 & (62.50\%)\\
Moderately Difficult to Obtain & 16 & 9 & 25 & (36.00\%)\\
Very Difficult to Obtain & 9 & 5 & 14 & (35.71\%)\\
\midrule
\addlinespace
Any restriction & 39 & 26 & 65 & (40.00\%)\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{*} Percentages are calculated as the number of 'Yes' divided by the}\\
\multicolumn{5}{l}{total number of responses. An article can have multiple categories}\\
\multicolumn{5}{l}{of data; the sum of responses is therefore higher than the number}\\
\multicolumn{5}{l}{of articles.}\\
\end{tabular}
\end{table}