### 2025-07-13 outputs/stata_downloads_regional_stats_2024_nochina.tex

% latex table generated in R 4.5.1 by xtable 1.8-4 package
% Fri Jun 27 19:46:23 2025
\begin{table}[ht]
\centering
\begin{tabular}{rlrrr}
  \hline
 & custom\_region & countries & Regional Downloads & Fraction \\ 
  \hline
1 & North America &   3 & 3509194 & 66.58 \\ 
  2 & Europe &  40 & 912383 & 17.31 \\ 
  3 & Rest of Asia &  45 & 521557 & 9.90 \\ 
  4 & Africa &  51 & 168663 & 3.20 \\ 
  5 & Latin America \& Caribbean &  28 & 127078 & 2.41 \\ 
  6 & Australia &   1 & 26703 & 0.51 \\ 
  7 & Rest of Oceania &   5 & 4763 & 0.09 \\ 
   \hline
\end{tabular}
\caption{Regional Statistics for Stata Downloads in 2024 (_nochina)} 
\label{tab:stata_downloads_regional_stats_2024_nochina}
\end{table}