### 2025-07-17 outputs/table_stata-licenses.tex

% latex table generated in R 4.5.1 by xtable 1.8-4 package
% Sat Jun 28 13:39:13 2025
\begin{table}[h]
\centering
\caption{Software licensing internationally} 
\label{tab:licensecost}
\begin{tabular}{lrrrrr}
  \toprule
Country & Average.monthly.income & Stata.license & Laptop & Percent.license & Percent.laptop \\ 
  \midrule
USA & 6,704 & 690 & 849 & 10.3 & 12.7 \\ 
  Greece & 1,883 & 690 & 934.8 & 36.6 & 49.6 \\ 
  Vietnam & 343 & 220 & 706.6 & 64.1 & 206 \\ 
   \midrule 
\multicolumn{6}{l}{\textit{Notes:} Amounts in USD as of 2025. Source for Stata license prices: Stata website.}\\
\multicolumn{6}{l}{ Sources for laptop prices: HP.com US and German websites, Shopee.vn for Vietnam.}\\
 \bottomrule
\end{tabular}
\end{table}